---
title: "Week-3"
author: "Denny Shin"
date: "2018-08-06"
output: workflowr::wflow_html
---

## **Working with csv**

To load data

```{r}
tasmania <- read.csv("data/tasmania.csv")
```

You can click on the file in the 'environment' tab to view the file.

You can now access tasmania.csv using array notation and built in functions
```{r}
dim(tasmania) # check data dimension
names(tasmania[, 1:5]) # names of first 5 columns
year <- tasmania[,1] # vector for year
s1 <- tasmania[,2] # vector for Burnie
s2 <- tasmania[,3] # vector for Cape Grim
```

We can start doing analysis!

```{r}
sd(s1) # standard deviation (of my sample)
IQR(s1) # interquartile range
quantile(s1, type=6)

sd(s2, na.rm=TRUE) # this removes missing values, NA
```

```{r}
hist(s1, freq=FALSE, xlab="Extreme rainfall (Burnie, Tas", col=8)
smooth.density = density(s1) # fits a smooth curve
lines(smooth.density, lty=2, lwd=2, col=2)
```

In hist() freq=FALSE makes the histogram to show density instead of count. You may also add nclass = or breaks = to control the number of bins.

```{r}
hist(s1, freq=FALSE, col=8, breaks=15)
```

boxplot:

```{r}
boxplot(s1, s2, horizontal = TRUE, names = c("Burnie Is", "Cape Grim"), col=c("yellow", "orange"))
```

empirical cdfs:

```{r}
ecdf1 <- ecdf(s1)
ecdf2 <- ecdf(s2)
plot(ecdf1)
plot(ecdf2, col=2, add=TRUE)
```

qqnorm:

```{r}
qqnorm(s1, main = "Normal QQ plot for S1")
qqline(s1)
```

there is more. do them later

## **Distributions**

## **Maximum likelihood estimation**

Here we use the MASS library and fitdistr function. We are finding the maximum likelihood numerically as many distributions, such as the Gumbel (EV) distribution has no closed form solution.

## **Method of Moments estimation**



## **Exercises**

##### **Q1**

Find Marrawah, Low Head and Dover:

```{r, eval=FALSE}
names(tasmania[,1:55])
```

Marrawah is at 7, Low Head is at 9, Dover is at 30.

```{r}
sM <- tasmania[,7]
sLH <- tasmania[,9]
sD <- tasmania[,30]

summary(sM)
summary(sLH)
summary(sD)

sd(sM)
sd(sLH, na.rm=TRUE)
sd(sD)

boxplot(sM, sLH, sD, horizontal=TRUE, names=c("Marrawah", "Low Head", "Dover"), col=c("yellow", "orange", "blue"))
```

#### **Q2**

I consult the help(boxplot) page.




