---
title: "Week-2"
author: "Denny Shin"
date: "2018-07-30"
output: workflowr::wflow_html
---

## **Vectors and Matrices**

To create a basic vector
```{r}
x <- c(4.1, -1.3, 2.5, -.6, -21.7)
x
```

You can combine vectors
```{r}
x <- c(2, 3)
y <- c(5, 7)
z <- c(x, y)
z
```

R will always do operations on vectors element by element
```{r}
c(1,2) + c(4,5)
1:5 + 3
```


Elements in vectors can called by [i]

We can also do ranges
```{r}
n <- 5
1:n
1:n-1
1:(n-1)
```

Sequences can be created by
```{r}
seq(1, 2, by=0.1)
seq(1, 2, length.out=20)
```

Matrices can be created by
```{r}
A <- matrix(c(2, -1, 4, 2, 1, 5), nrow=2)
A
```

Matrix operations are
```{r}
A[1,2]
A[1, ] # first row
mean(A)
```

## **Characters**

Characters must be typed in ""

More elaborate characters can be written with the paste() function
```{r}
paste(c("X", "Y"), 1:10, sep="")
```
It will paste the character with a number seperated by sep=""

## **Funtioncs**

There are many standard functions built into R.

rnorm creates random samples from a normal distribution
```{r}
x <- rnorm(10, mean=10, sd=2)
x
```

## **Exercises**

#### Q1
```{r}
sum(log(1:100))
```

#### Q2
```{r}
x <- rnorm(100, mean=1, sd=2)
mean(x^2)
```

#### Q3
qnorm is the normal distribution but with certain presets.

qnorm takes in vector of probabilities, p, not the number of observations

qnorm sets lower.tail=TRUE and log.p=FALSE

lower.tail=TRUE means probabilities are P[X<=x]
log.p=TRUE mean probabilities  are given as log(p)
```{r}
qnorm(0.1)
qnorm(0.1, lower.tail = FALSE)

a = 0.9
qnorm(a, lower.tail = FALSE)
```

#### Q4
exponential function with mean=1
```{r}
exp1pdf <- function(x){
  if (x>=0){
    return(exp(-x))
  } else {
    return(0)
  }
  
}
```

Comparing with dexp()

```{r}
exp1pdf(1) # should give same value as dexp(1)
dexp(1)
```

Also checking
```{r}
exp1pdf(-1) #should give 0
dexp(-1)
```

#### Q5



