---
title: "Week-2"
author: "Denny Shin"
date: "2018-07-30"
output: workflowr::wflow_html
---

## **Vectors and Matrices**

To create a basic vector
```{r}
x <- c(4.1, -1.3, 2.5, -.6, -21.7)
x
```

You can combine vectors
```{r}
x <- c(2, 3)
y <- c(5, 7)
z <- c(x, y)
z
```

R will always do operations on vectors element by element
```{r}
c(1,2) + c(4,5)
1:5 + 3
```


Elements in vectors can called by [i]

We can also do ranges
```{r}
n <- 5
1:n
1:n-1
1:(n-1)
```

Sequences can be created by
```{r}
seq(1, 2, by=0.1)
seq(1, 2, length.out=20)
```

Matrices can be created by
```{r}
A <- matrix(c(2, -1, 4, 2, 1, 5), nrow=2)
A
```

Matrix operations are
```{r}
A[1,2]
A[1, ] # first row
mean(A)
```

## **Characters**

Characters must be typed in ""

More elaborate characters can be written with the paste() function
```{r}
paste(c("X", "Y"), 1:10, sep="")
```
It will paste the character with a number seperated by sep=""

## **Functions**

There are many standard functions built into R.

rnorm creates random samples from a normal distribution
```{r}
x <- rnorm(10, mean=10, sd=2)
x
```

## **Exercises**

#### **Q1**
```{r}
sum(log(1:100))
```

#### **Q2**
```{r}
x <- rnorm(100, mean=1, sd=2)
mean(x^2)
```

#### **Q3**
qnorm is the inverse cdf of normal distribution.

qnorm takes in vector of probabilities, p, not the number of observations.

qnorm sets default lower.tail=TRUE and log.p=FALSE

lower.tail=TRUE means probabilities are P[X<=x]
log.p=TRUE mean probabilities  are given as log(p)
```{r}
qnorm(0.1)
qnorm(0.1, lower.tail = FALSE)

a = 0.9
qnorm(a, lower.tail = FALSE)
```

#### **Q4**
exponential function with mean=1
```{r}
exp1pdf <- function(x){
  if (x>=0){
    return(exp(-x))
  } else {
    return(0)
  }
  
}
```

Comparing with dexp()

```{r}
exp1pdf(1) # should give same value as dexp(1)
dexp(1)
```

Also checking
```{r}
exp1pdf(-1) #should give 0
dexp(-1)
```

#### **Q5**

```{r}
data = c(10.39, 10.43, 9.99, 11.17, 8.91, 11.20, 11.38, 7.74, 10.61, 11.11)
summary(data)

min(data)

boxplot(data, horizontal = TRUE)
```

looking at the boxplot, there is one outlier 7.74

#### **Q6**

```{r}
data = c(9.9, 4.7, 20.5, 1.8, 4.7, 9.8, 20.5, 20.2, 6.5, 3.0)

hist(data, freq=FALSE, xlab="Histogram of daily maximum rainfall (cm) in Tasmania", col=8)

qqnorm(data)
qqline(data)
```
plotting the normal QQ-plot is not good enough. We know that this distribution is more likely to follow the extreme value (EV) distribution.

Let's call our rainfall random variable $Y$. We can then say that $Y=\theta+\xi Z$ where $Z$ is the standard EV dist. with cdf $F(z)=exp(-exp(-z))$ and inverse cdf $F^{-1}(z)=-ln(-ln(p))$.

Using the QQplot method:

```{r}
n=10
p=1:n/(10+1) # p=k/(n+1)
data.sorted = sort(data) # order statistics
Finv = -log(-log(p)) # inverse cdf for EV dist.

plot(Finv, data.sorted, pch=19,
     xlab="EV quantiles",
     ylab="Sample quantiles")

# find line of best fit
# he says don't worry too much about these, they will be explored in later labs
fit = lm(data.sorted ~ Finv)
abline(fit)

```

To find $\theta$ and $\xi$ we look at the intercept and gradient:

```{r}
coef(fit)
```

#### **Q7**

To simulate $X_i$ we need the inverse cdf: $F^{-1}(p)=(2p-1)^{1/3}$. We then simutale $Y$ to calculate $Pr(-0.3<Y<1.5)$.

```{r}

```





